---
title: "Very basics of R coding key"
author: "John Shorter"
date: "Feb/14/2024"
output:
  html_document:
    toc: yes  
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
```

# Exercise

Today you have learned some basic syntax of R. Now it's time for you to practice.

## Q1:

# Exercise

Today you have learned some basic syntax of R.
Now it's time for you to practice.

## Q1:

1.  Insert a new code chunk
2.  Make this matrix

| 1  | 1 | 2 | 2 |
| 2  | 2 | 1 | 2 |
| 2  | 3 | 3 | 4 |
| 1  | 2 | 3 | 4 |

and save it as an item called `my_mat2`.

3. Select the 1st and 3rd rows and the 1st, 2nd and 4th columns, and save it as an item.

4. Take the square root for each member of my_mat2, then take log2(), and lastly find the maximum value.
Use the pipe syntax. The command for maximum is `max()`.



```{r}
my_mat2 <- rbind(
  c(1, 1, 2, 2),
  c(2, 2, 1, 2),
  c(2, 3, 3, 4),
  c(1, 2, 3, 4)
)

my_mat3 <- my_mat2[c(1,3), -3]
my_mat3

my_mat2 %>% 
  sqrt() %>% 
  log2() %>% 
  max()

```

## Q2:

1.  Use the following info to make a data frame and save it as an item called "people".
    Anne 25 years old, Bert is 83 years old, and Chad is 51 years old.
    Their hair colors are black, brown, and blonde, respectively.
    (Hint: How many columns do you have to have?)

2. Pull out the column with the ages.
    Use the `$` syntax.


```{r}
people <- data.frame(
  name = c("Anne", "Bert", "Chad"),
  age = c(25, 83, 51),
  hair = c("Black", "Brown", "Blonde")
) 

people

people$age
```
