---
title: "Formatting ggplots key"
author: "John Shorter"
date: "Feb/21/2024"
output:
  html_document:
    toc: yes  
  html_notebook:   
    number_sections: yes    
    toc: yes  
    toc_float: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

You might need to install some of these packages

```{r}
library(ggplot2) 
library(tidyr)
library(dplyr)
library(readr)
library(readxl)
```

# Exercise - Now time for you to try

Let's use the R built-in dataset CO2 as example: 
Ref: Potvin, C., Lechowicz, M. J. and Tardif, S. (1990), Ecology, 71, 1389--1400.

```{r}
head(CO2)
str(CO2)
```

Task: produce a plot that includes the info for Type and Treatment, with uptake on Y axis.
Let's say we are interested in the effect of "Treatment" in each Type.

Hint:

1.  What should be on the x axis?
2.  What variable do you color or fill with?
3.  What faceting layout do you want to use?

Now try to make the best plot you can!


```{r}
CO2 %>% 
  ggplot(aes(x = Treatment, y = uptake)) +
  facet_grid(. ~ Type, switch = "x") + 
  geom_point(aes(fill = Treatment), size = 3, color = "white", shape = 21,
             position = position_jitter(0.1, seed = 666)) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2) +
  scale_fill_manual(values = c("burlywood3", "slateblue3"),
                    limits = c("nonchilled", "chilled")) +
  labs(x = "Type") +
  theme_minimal()+
  theme(legend.position = "top",
        axis.line = element_line(size = 1.2),
        panel.spacing.x = unit(1, "lines"),
        text = element_text(size = 12, color = "black", face = "bold"),
        axis.text.x = element_blank(),
        axis.text = element_text(size = 12, color = "black", face = "bold"),
        strip.placement = "outside"
        )
```

# Exercise 2 - Something fun!

Let's use the Pokemon data set from https://www.kaggle.com/datasets/abcsds/pokemon?resource=download.

```{r}
Pokemon <- read.csv("/work/Data/Rmodule_data/Pokemon.csv", stringsAsFactors = FALSE)

Pokemon2 <- subset(Pokemon, Type.1 =='Water' | Type.1 =='Fire'| Type.1 =='Grass')
Pokemon2 <- subset(Pokemon2, Generation =='1' | Generation == '2')
#Don't worry about subsetting for now, we will go over this in a later lesson.
```


```{r}
head(Pokemon2)
str(Pokemon2)
```

Task: produce a plot that compares speed of fire, grass, and water pokemon from generations 1 and 2. Let's keep it simple and use only the primary type. (Type.1 column)

Hint:

1.  What should be on the x axis?
2.  What variable do you color or fill with?
3.  What faceting layout do you want to use?

Now try to make the best plot you can!

Which type is the fastest?
Which pokemon is the fastest?

```{r}
Pokemon2 %>% 
  ggplot(aes(x = Type.1, y = Speed)) +
  facet_grid(. ~ Generation, switch = "x") + 
  geom_point(aes(fill = Type.1), size = 3, color = "white", shape = 21,
             position = position_jitter(0.1, seed = 666)) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2) +
  scale_fill_manual(values = c("red3", "slateblue3", "green3"),
                    limits = c("Fire", "Water", "Grass")) +
  labs(x = "Generation") +
  theme_bw()+
  theme(legend.position = "top",
        axis.line = element_line(size = 1.2),
        panel.spacing.x = unit(1, "lines"),
        text = element_text(size = 12, color = "black", face = "bold"),
        axis.text.x = element_blank(),
        axis.text = element_text(size = 12, color = "black", face = "bold"),
        strip.placement = "outside"
        )
```