---
title: "Analyses of proportions, contingency tables and enrichment key"
author: "John Shorter"
date: "Mar/20/2024"
output:
  html_document:
    toc: yes  
  html_notebook:   
    number_sections: yes    
    toc: yes  
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2) 
library(tidyr)
library(dplyr)
library(readr)
library(readxl)
library(RColorBrewer)
library(viridis)
```

# Exercise

## Chi-squared goodness of fit test

You did two crosses. 
You crossed a Aa mother to a AA father, you got 30 Aa and 60 AA. 
You crossed a AA mother to a Aa father, you got 55 Aa and 50 AA. 
(The expected proportions should be 1:1.)

Is there distortion of segregation ratio in either of the crosses? 
If so, which parent is the defect associated with?

```{r}
chisq.test(x = c(30, 60), p = c(0.5, 0.5))
chisq.test(x = c(55, 50), p = c(0.5, 0.5))
```

There is a significant distortion of segregation ratio in the Aa mom x AA dad cross (p = 0.0016) 
This indicates the defects are associated with the maternal parent.



Now let's look at the built in dataset diamonds, and see if there is a difference in the cut and color variables.

```{r}
# Load the diamonds dataset
data(diamonds)

# View the first few rows of the dataset
head(diamonds)

```


```{r}
# Plot the frequency of cut and color
ggplot(diamonds, aes(x = cut, fill = color)) +
  geom_bar(position = "fill") +
  labs(title = "Frequency of Cut by Color",
       x = "Cut",
       y = "Proportion",
       fill = "Color") +
  theme_minimal()


# Perform Chi-square test
chisq_result <- chisq.test(diamonds$cut, diamonds$color)

# Print the result
print(chisq_result)


```



